---
title: "Rayleigh distillation - example of numerical and analytical solution"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document: 
    df_print: paged
    toc: yes
    toc_float: true
    code_folding: show
editor_options:
  chunk_output_type: inline
---

```{r "setup", warning=FALSE, message=FALSE}
# make sure to install this package for solving ordinary differential equations
library(deSolve) 
library(tidyverse)
library(tibble)
```

# Define the system

To solve an ODE system, one needs to define a function that represents the differential equations. Here we are considering the irreversible reaction $A \rightarrow B$ in a closed system and are keeping track of our light ($L$) and heavy ($H$) isotope pool in both the starting reactant reservoir ($A$) and product pool $B$. As discussed in class this could, this could also be employed to describe an equilibrium reaction where the product is lost (e.g. cloud condensation and precipitation).

Assuming some constants $k_H$ and $k_L$ that describe the different reaction rates for the heavy and light isotope (and define the fractionation factor $\alpha$ of the system), the four variable pools ($L$ and $H$ isotopes in both $A$ and $B$) can be described how:

## Derivatives function

```{r}
calc_derivs <- function(t, y, parms) {
  with (c(as.list(y), parms), {
    
    dAH <- 
    dBH <- 
    
    dAL <- 
    dBL <- 
    
    # make sure to return in the same order as initial values
    return(list(c(dAL, dAH, dBL, dBH)))
  })
}
```

## Variables

```{r}
# define time sequence (how many time steps to run)
# the meaning of time depends on the value of the rate constants (if rate constants are large,
# small time steps will lead to big changes of the system and vice versa)
times <- seq(0, 100, by = 0.1)

# define starting values of the system
# these are also arbitrary isotope ratios, 
# change the values to see what happens
variables_initial <- c(
  L_A = ,
  H_A = ,
  L_B = ,
  H_B = 
)
variables_initial
```

## Parameters

```{r}
# define additional parameters for this system
# here I'm showing how one could use multiple scenarios to explore
# the consequences of parameter changes (using the tribble funtion
# to define a data frame row by row)
parameters <- 
  tribble(
    # scenario name, alpha (heavy over light rates ratio), light reaction rate
    ~scenario,    ~alpha_HoL, ~k_L,
    "Enrichment", 1.005,      0.1,
    "Depletion",  0.995,      0.1
  ) %>% 
  # derived parameters
  mutate(
    k_H = k_L * alpha_HoL
  )
parameters
```

# Solve the model

```{r, warning=FALSE}
# solve the model for each parameter set
out <-
  parameters %>% 
  # group by scenario
  group_by(scenario) %>% 
  # do the calculation for each parameter set
  do({
    ode(
      y = variables_initial, 
      times = times, 
      func = calc_derivs, 
      parms = as.list(.)) %>% 
      as.data.frame()
  })

# see what ode solver returns
out
```

# Evaluation

```{r}
# merge parameters into the results

# calculate ratios, deltas and mass fractions

# calculate expected values from theoretical considerations

# calculate the approximate delta values from theoretical considerations

```

# Plotting

Compare the different ratios and alphas.

 - How does the numerical simlulation compare with the analytical solution?

 - How do the delta approximation equations compare with the exact result from the numerial simulation? How might this change with stronger isotope fractionation?

 - The fractionation factors are only about 5 permil (either towards depleting or enriching the product). What do you notice about the range of isotope values for the starting reservoir and accumulated product?

```{r}

```

